Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-29 manuscript.Rmd
Edit Discussion

A number of other metrics provide insight not only into the amount of structure
between the systems but about the nature of this shared structure.
The average length of the line structures is called *mean line* (L), which can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
*max line* (*maxL*); it indicates the longest instance of being “stuck” in a
particular state together. The total *number of lines* on the plot (*NRLINE*) 
scales with structure: The more lines that exist, the more coupled the systems are. The
complexity or variability of line lengths is captured by a metric called
*entropy* (*ENTR*)---which may also be normalized by the total number of lines
(i.e., *normalized entropy*; *rENTR*)---and indicates
the stability and predictability of the system. When one of the signals gets
“stuck” in a state and the other moves on to other states, this results in
vertical line structures, which are quantified in *laminarity* (*LAM*;
i.e., the proportion of points forming vertical line structures) 
and trapping time (*TT*; i.e., the average length of the
vertical trajectories). For more detail on CRQA metrics, see @R-crqa.

@@

# Discussion
In the present study, we investigated the dynamics of real-world events
and online communication on Twitter in Syria during Arab Spring. We examined
how social mobilization both virtually and physically evolve over time in
tandem. Prior to now, the extent to which virtual mobilization on social
media platforms such as Twitter has been largely unexplored. Using the
nonlinear methods cross-recurrence quantification analysis and windowed
cross-recurrence quantification analysis, we investigated the theory that
virtual mobilization is increasingly reflective of real-world events
during times of strife.

## Cross-Recurrence Quantification Analysis
### Syria as Target
The significant result of DET for social cohesion and both the all events
and negative events time series confirms our hypothesis, that increased
salience of events is reflected in social cohesion on Twitter. The frequency
of all events and of negative events thus result in a significant number
of shared trajectories that are statistically different from chance. The
non-significant results for all metrics for positive event count also
confirm our hypothesis, and gives support to the theory that positive events
are less coupled to social cohesion than the more salient, negative events
during times of turmoil. The inevitable increased salience of event
frequency and event violence thus leads to increased coupling with Twitter
social cohesion.

### Syria as Source and Target
By adding Syria as a source in the data filtering, the dataset increased
in size by 1,690 samples. These additional observations are times where
Syria initiated an event with another country. Events initiated by Syria
but directed internally within the country were already captured by the
target only filtering. The increase in coupling between positive events
and social cohesion is thus due to the inclusion of these externally
directed events. Though CRQA metrics for count of all and count of
negative events with social cohesion were largely unaffected, count of
positive events and social cohesion showed significant levels of
determinism, number of lines, max line length, and laminarity that were
not present in the target filtered data. Given that the Syria during
this time was still under Authoritarian rule, the government may have
been initiating more positive external events to prevent intervention
by other countries or try to cover up what was happening internally.
We believe that the increased coupling of positive events is thus not
reflective of actual social cohesion, but happened by chance. The
inflation of the number of positive events when including source
observations is causing increased coupling, but it is more likely that
the Twitter users were using the platform to raise awareness of the
violent, internal events that were being inflicted on them by their
government officials and armies and not reporting on the positive
external events that did not directly affect them or their safety.
Further research should explore the semantic content of this tweet
corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs;
Figures \@ref(fig:plot-rp-targ-all)-\@ref(fig:plot-rp-source-targ-neg)),
there appears to be a shift in the dynamics of these systems around
observation 35, or May 3rd, 2012. To further investigate these local
dynamics, we performed a sliding window technique known as windowed
cross-recurrence quantification analysis (WCRQA).

## Windowed-Recurrence Quantification Analysis
### Syria as Target
The time period of differing dynamics, from observation 38 through 43,
appears to be captured by WCRQA of social cohesion and count of negative
events for both RR and DET (Figure \@ref(fig:plot-wcrqa-targ-all)). The
amount of recurrence and determinism for these periods almost consistently
lie above chance, suggesting that the intensity of negative events coincide
with the degree of social cohesion found on Twitter, and that these states
carry across multiple days in a way that is different from that of chance.

The above-chance levels of determinism and recurrence rate are demonstrating
the coupling of social cohesion and counts of all and negative events during
times of strife. We can see this string of highly violent events reflected
in the mode ICEWS event for each day (Figure \@ref(fig:plot-mode-event)),
where from May 1st, 2020 to May 15th, 2020, the most frequently occurring
event was almost consistently -10, the most violent event in the coding
system. These dates directly correlate with the above chance RR and DET
metrics for the WCRQA of social cohesion and count of negative events.
Furthermore, for WCRQA of social cohesion and positive event count during
this time period, DET was entirely within chance (Figure
\@ref(fig:plot-wcrqa-targ-all)). This supports our hypothesis, as the
positive events are seemingly not coupled with the social cohesion on Twitter
during civil uprisings during the Arab Spring.

### Syria as Source and Target
WCRQA of recurrence rate on the source and target filtered data remained
largely the same for count of all events and social cohesion, suggesting
that large numbers on a given day will be reflective of increased levels
of social cohesion. Determinism showed more frequent above-chance metrics
for windows 39 and up, suggesting that the inclusion of externally
directed events lead to higher event counts during these days and then
was more reflective of the level of social cohesion on Twitter. WCRQA for
count of positive events and social cohesion remained largely the same for
RR and DET as the source only data. This again seems to support the idea
that these two aren't actually coupled, but rather that the inflation of
the positive counts lead to a spurious relationship. Count of negative
events and social cohesion also showed similar trends for RR and DET,
again attesting to the theory that negative events were coupled to social
cohesion on Twitter while positive events happened by chance.

## Conclusion
Considering the invasiveness of social media in our everyday lives, it is
inevitable that individuals involved in real-world social mobilization have
utilized these platforms to their advantage. If consistent across other civil
wars, these results have serious implications for monitoring global
movements. While traditional media in regions of authoritarian rule are
often biased, the self-organized nature of Twitter poses a new avenue to gain
insight into truths about up-risings and social mobilization. By
conceptualizing Twitter and real-world events as two components of an
inextricably linked dynamical system, the levels of social cohesion within the
virtual platform could be monitored for fluctuations indicating global shifts
in power and peace. Social media mobilization can thus help alert us to
real-world mobilization not communicated in mainstream media and allow for 
faster intervention.

>>>>

A number of other metrics provide insight not only into the amount of structure
between the systems but about the nature of this shared structure.
The average length of the line structures is called *mean line* (L), which can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
*max line* (*maxL*); it indicates the longest instance of being “stuck” in a
particular state together and is a measure of the stability of the system. 
The total *number of lines* on the plot (*NRLINE*) 
scales with structure: The more lines that exist, the more coupled the systems are. The
complexity or variability of line lengths is captured by a metric called
*entropy* (*ENTR*)---which may also be normalized by the total number of lines
(i.e., *normalized entropy*; *rENTR*)---and indicates
the stability and predictability of the system. When one of the signals gets
“stuck” in a state and the other moves on to other states, this results in
vertical line structures, which are quantified in *laminarity* (*LAM*;
i.e., the proportion of points forming vertical line structures) 
and trapping time (*TT*; i.e., the average length of the
vertical trajectories). For more detail on CRQA metrics, see @R-crqa.

@@

# Discussion
In the present study, we investigated the dynamics of real-world events
and online communication on Twitter in Syria during Arab Spring. We examined
how social mobilization both virtually and physically evolve over time in
tandem. Using the
nonlinear dynamical time series analyses, we investigated the theory that
virtual mobilization is increasingly reflective of real-world events
during times of strife.

## Cross-Recurrence Quantification Analysis
### Syria as Target
The significant result of DET for social cohesion with all and negative 
events supports our hypothesis that
event salience is coupled with online social cohesion The frequency
of all and negative events are thus associated with greater
shared trajectories with social cohesion that would be expected by chance. 
However, when we specifically target positive events, the connection
to social cohesion fades, providing further
support for our hypothesis. Positive events
may be less coupled to social cohesion than more salient negative events
during times of turmoil; by contrast, overall event frequency and
negative event frequency are coupled with Twitter social cohesion.

### Syria as Source and Target
By adding Syria as a source in the data filtering, the dataset increased
in size by 1,690 samples. These additional observations are times where
Syria initiated an event with another country. Events initiated by Syria
but directed internally within the country were already captured by the
target-only filtering. The increase in coupling between positive events
and social cohesion can thus only be due to the externally
directed events. Though the coupling between social cohesion with
negative and all events were largely unchanged, 
positive events and social cohesion showed greater stability
in their coupling (reflected in maxL); while TT was no longer
significant in these analyses, LAM (related to TT) did
reach statistical significance.

While interesting, the underlying meaning or drivers of this change
in results requires further investigation that incorporates the
actual content of the tweets themselves. One possible reason (among
many) may be that the finding was spurious. The positive external 
events from Syria may have been self-serving, aiming to prevent intervention
by other countries or to cover up what was happening internally.
During these moments, the increased cohesion may reflect a surge from 
the Syrian people trying to raise awareness of the situation or
to decry the Syrian government's attempts to improve their international
standing. In this case, the increased coupling of positive events would
coupled not with shared positive affect in social media but instead
to shared negative affect in social media in response. However, we
cannot test this hypothesis with the current data; future  
research should explore the semantic content of this tweet
corpus to determine the nature of the socially cohered language.

## Windowed-Recurrence Quantification Analysis

### Syria as Target
The time period of differing dynamics---from observation 38 through 
43---appears to be captured by the RR and DET values in
WCRQA of social cohesion and count of negative
events (Figure \@ref(fig:plot-wcrqa-targ-all)). The
amount of recurrence and determinism for these periods almost consistently
lie above chance, suggesting that the intensity of negative events coincide
with the degree of social cohesion found on Twitter and that these 
shared trajectories extend across multiple days.

The above-chance levels of determinism and recurrence rate demonstrate
the coupling of social cohesion to both total and negative events during
times of strife. During this period, 
the modal ICEWS event for each day (Figure \@ref(fig:plot-mode-event))
during May 1st---15th was almost consistently -10, the most violent 
event category. These dates directly correlate with the above-chance RR and DET
metrics for the WCRQA of social cohesion and count of negative events.
Furthermore, for WCRQA of social cohesion and positive event count during
this time period, DET was entirely within chance (Figure
\@ref(fig:plot-wcrqa-targ-all)). This supports our hypothesis, as the
positive events are seemingly not coupled with the social cohesion on Twitter
during civil uprisings during the Arab Spring.

### Syria as Source and Target
WCRQA of recurrence rate on the source and target filtered data remained
largely the same for count of all events and social cohesion, suggesting
that large numbers on a given day will be reflective of increased levels
of social cohesion. DET showed more frequent above-chance metrics
for windows 39 and up, suggesting that the inclusion of externally
directed events lead to higher event counts during these days and then
was more reflective of the level of social cohesion on Twitter.

WCRQA metrics for positive-only and negative-only event counts with 
social cohesion remained largely the same for
RR and DET as the source-only data. This may support the ideas suggested
in the CRQA section, but again, only future work including the *content*
of those tweets can directly speak to that explanation.

## Conclusion
Given the pervasiveness of social media, it is unsurprising
that social activists and protesters have
utilized these platforms to support real-world mobilization. 
Here, we quantitatively explored the connection
between online social cohesion and real-world events. In addition
to introducing political scientists to a family of powerful nonlinear 
dynamics analyses, our results have serious implications for monitoring global
movements. While traditional media in regions of authoritarian rule are
often biased, the self-organized nature of Twitter poses a new avenue to gain
insight into truths about uprisings and social mobilization. By
conceptualizing Twitter and real-world events as two elements of an
inextricably linked dynamical system, the levels of social cohesion within 
social media platforms could be monitored for fluctuations indicating shifts
in power and peace. Social media mobilization may thus help alert the world to
citizens' actions and reactions that are suppressed by government regimes,
providing insight into events that may not otherwise be broadcast.
